|

### 2022-09-16 article/article.Rmd
creating tex version for publication

---
title:  |
        | The Effect of Legislature Size on Public Spending:
        | A Meta-Analysis
author:
- "Danilo Freire[^danilo]"
- "Umberto Mignozzetti[^umberto]"
- "Catarina Roman[^catarina]"
- "Huzeyfe Alptekin[^huzeyfe]"
date: \today
thanks: "Freire, Mignozzetti, and Roman contributed equally to all parts of the manuscript. Alptekin helped collect data in the early stages of the project. The authors thank Guilherme Duarte, Robert Myles McDonnell, and David Skarbek for their constructive feedback. We also thank Cedric Antunes, Luis Castro, Giovanna França, Julia Oriente, and Lucas Mingardi for their excellent research assistance. Replication materials are available at <https://github.com/danilofreire/legislature-size-meta-analysis>. We kindly acknowledge funding from the São Paulo State Science Foundation (FAPESP grant number 2018/00646-1)."
abstract: "In a seminal article, @weingast1981political argue that there is a positive relationship between legislature size and inefficiency in public expenditures. Their proposition is currently known as the \"law of $1/n$\" and has been widely debated in political science and public administration. However, recent studies have questioned the validity of the theory. In this letter, we conduct the first meta-analysis that assesses the generality of the \"law of $1/n$\". Based on a sample of 30 articles, we find no robust evidence suggesting that legislature size has either a positive or a negative effect on government budgets. Yet the aggregate results mask considerable heterogeneity. Our findings provide moderate support for the \"law of $1/n$\" in unicameral legislatures and in upper houses, but they also indicate that papers using panel/fixed-effects models or regression discontinuity designs report negative public spending estimates. We find only limited evidence that electoral systems impact public spending, which suggests that proportional representation systems may not be more prone to overspending than majoritarian ones."
abstractspacing: double
keywords: distributive politics; law of $1/n$; legislature size; meta-analysis; public spending
jelcodes: H21; H23; H50; H61
fontsize: 11pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

@@

@@

\nocite{baskaran2013coalition, bradbury2009spatially, drew2017price,
erler2007termlimits, fiorino2007legislature, hohmann2017effect,
kessler2014communication, lewis2019legislature, lledo2003electoral,
mukherjee2003politicalparties, petterssonlidbom2012size, schaltegger2009large,
stein1998institutional, mukherjee2003politicalparties, macdonald2008impact,
matsusaka2005endogeneity, ricciuti2004legislatures, coate2011government}

>>>>

---
title:  |
        | The Effect of Legislature Size on Public Spending:
        | A Meta-Analysis
author:
- "Danilo Freire[^danilo]"
- "Umberto Mignozzetti[^umberto]"
- "Catarina Roman[^catarina]"
- "Huzeyfe Alptekin[^huzeyfe]"
date: \today
thanks: "Freire, Mignozzetti, and Roman contributed equally to all parts of the manuscript. Alptekin helped collect data in the early stages of the project. The authors thank Guilherme Duarte, Robert Myles McDonnell, and David Skarbek for their constructive feedback. We also thank Cedric Antunes, Luis Castro, Giovanna França, Julia Oriente, and Lucas Mingardi for their excellent research assistance. Replication materials are available at <https://github.com/danilofreire/legislature-size-meta-analysis>. We kindly acknowledge funding from the São Paulo State Science Foundation (FAPESP grant number 2018/00646-1)."
abstract: "In a seminal article, @weingast1981political argue that there is a positive relationship between legislature size and inefficiency in public expenditures. Their proposition is currently known as the \"law of $1/n$\" and has been widely debated in political science and public administration. However, recent studies have questioned the validity of the theory. In this letter, we conduct the first meta-analysis that assesses the generality of the \"law of $1/n$\". Based on a sample of 30 articles, we find no robust evidence suggesting that legislature size has either a positive or a negative effect on government budgets. Yet the aggregate results mask considerable heterogeneity. Our findings provide moderate support for the \"law of $1/n$\" in unicameral legislatures and in upper houses, but they also indicate that papers using panel/fixed-effects models or regression discontinuity designs report negative public spending estimates. We find only limited evidence that electoral systems impact public spending, which suggests that proportional representation systems may not be more prone to overspending than majoritarian ones."
abstractspacing: double
keywords: distributive politics; law of $1/n$; legislature size; meta-analysis; public spending
jelcodes: H21; H23; H50; H61
fontsize: 11pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: yes
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

@@

# Backmatter Headings

* **Supplementary Materials**: The supplementary materials for this paper can be found at <https://URL_JOURNAL_HERE>.

* **Data Availability Statement**: Replication data for this paper can be found at <https://doi.org/10.7910/DVN/5DEGYP>. Replication materials are also available at <https://github.com/danilofreire/legislature-size-meta-analysis>.

* **Acknowledgements**: The authors thank Guilherme Duarte, Cathy Hafer, Robert Myles McDonnell, David Skarbek, and Howard Rosenthal for their constructive feedback. We also thank Cedric Antunes, Luis Castro, Giovanna França, Julia Oriente, and Lucas Mingardi for their excellent research assistance.

* **Author Contributions**: Freire, Mignozzetti, and Roman contributed equally to all parts of the manuscript. Alptekin helped collect data in the early stages of the project.

* **Financial Support**: We kindly acknowledge funding from the São Paulo State Science Foundation (FAPESP), grant number 2018/00646-1.

* **Competing Interests**: None.

@@

\nocite{datasetCite, baskaran2013coalition, bradbury2009spatially, drew2017price,
erler2007termlimits, fiorino2007legislature, hohmann2017effect,
kessler2014communication, lewis2019legislature, lledo2003electoral,
mukherjee2003politicalparties, petterssonlidbom2012size, schaltegger2009large,
stein1998institutional, mukherjee2003politicalparties, macdonald2008impact,
matsusaka2005endogeneity, ricciuti2004legislatures, coate2011government}
